---
title: "Sensitivity_ASD_PTSD"
author: "Brian Holt"
date: "11/7/2021"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(gt)
library(tidyverse)
```

https://www.ptsd.va.gov/publications/rq_docs/V24N1.pdf  2013 update

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5381710/  says it's pretty good 2017 maybe just for kids?

well, midnight phone research led to me losing the cite.   Having kids makes it hard to work.  I either ignore them or the work.  

sensitivity was `r sen = .86`
specificity was `r spec =.29`

Now, sensitivity is found By taking the total cases marked as true positives / true positives plus false negatives 

or TP / (TP + FN)

While specificity is  TN / (TN + FP)

https://www.nottingham.ac.uk/nmp/sonet/rlos/ebp/sensitivity_specificity/page_four.html


So, using some algebra

```{r, grid set up, echo=T}
tp <- 9
tn <- 69
fp <- 1
fn <- 21
(ptsd.absent <- fp+tn)
(ptsd.present <- tp+fn)
(tot.pos <- tp+fp)
(tot.neg <- tn+fn)

tt <- tibble(metric =c('tp','fp','tn','fn'),
             value = c(tp,fp,tn,fn),
             column= c(1,2,2,1))
gt(tt) %>% 
   tab_header(
    title = "Sensitivity/Specificity example",
    subtitle = "ASD as a diagnostic test for PTSD"
  ) %>% 
  tab_row_group(
    label = 'test outcome',
    rows = c(1,4)
  )
```

.86 = Sensitivity = TP/TP+FN


```{r,echo=T}
p.ptsd <- rbinom(1e5,1,.03)
asd <- rep(0,1e5)
df<- data.frame(p.ptsd,asd)


df[p.ptsd==1,'asd'] <- rbinom(length(df[p.ptsd==1,'asd']),1,.85)
df[p.ptsd!=1,'asd'] <- rbinom(length(df[p.ptsd!=1,'asd']),1,.7)
(sens <- table(df)[4]/(table(df)[4]+table(df)[2]))
(spec <- table(df)[1]/(table(df)[1]+table(df)[3]))

'tp <- table(df)[4]
fp <- table(df)[3]
tn <- table(df)[1]
fn <- table(df)[2]
'
df %>% pivot_longer(c(p.ptsd,asd)) %>%
  group_by(name) %>% summarise(total = sum(value))
  ggplot(aes(y=value))+geom_histogram()+facet_grid(~name)




s.1<- sample(p,100,replace=T)


```





One thing might be to do the above, but to use a gradient of severity to give it continous data